---
title: "Reproducible Research Project"
output: html_document
---

First set the working directory:

```{r setup, include=FALSE}
setwd("C:/Users/akyani/Documents/Data Science/Fifth/Second Week/Project/Factivity/")

```
Then load Data
```{r}
data = read.csv ("activity.csv")
```
Here are the answers to each part of the project:

Histogram of the total number of steps taken each day
``` {r}
library(dplyr)
by_day = group_by(data,date)
total_by_day = summarise(by_day,total = sum(steps))
```
Here is the plot 
``` {r}
hist(total_by_day$total,main = "Total Steps in Each Day", xlab = "Steps")
```

Mean and median number of steps taken each day
```{r}
dataMean = summarise(by_day,mean = mean(steps,na.rm = TRUE))
dataMedian= summarise(by_day,median = median(steps,na.rm = TRUE))
```

Time series plot of the average number of steps taken

```{r}
by_interval = group_by(data,interval)
mean_by_interval = summarise(by_interval,mean = mean(steps,na.rm = TRUE))
plot(mean_by_interval$mean,type = "l",xlab = "Interval",ylab = "Average")
```

The 5-minute interval that, on average, contains the maximum number of step
```{r}
by_interval = group_by(data,interval)
max_by_interval = summarise(by_interval,max = max(steps,na.rm = TRUE))
```

Code to describe and show a strategy for imputing missing data
```{r}
NAData= sum(is.na(data))
```

Backup for data 
```{r}
imputed_data = data
```

Merge it to replace the NA values

```{r}
data3 = merge (data,mean_by_interval)
imputed_data$steps = ifelse(is.na(imputed_data$steps), data3$mean,data$steps)
```

Histogram of the total number of steps taken each day after missing values are imputed
```{r}

by_day_impute = group_by(imputed_data,date)
dataMean_impute = summarise(by_day,total = sum(steps,na.rm = TRUE))
hist(dataMean_impute$total,main = "Total Steps in Each Day", xlab = "Steps")
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
# Convert date to weekdays

imputed_data$day = weekdays(as.Date(imputed_data$date))

by_interval = imputed_data %>% group_by(interval,day) %>% summarise(average = mean(steps))

weekends = subset(by_interval,day %in% c("Saturday","Sunday"))
weekdays = subset(by_interval,!day %in% c("Saturday","Sunday"))


par(mfrow=c(2,1))
plot(weekends$average, type = "l",xlab = "Interval",ylab = "Average")
plot(weekdays$average, type = "l",xlab = "Interval",ylab = "Average")
```

